The table and figures below give linear fits of _S_-1 versus log(_N_). The table lists the number of samples, the intercept (b0), the slope (b1), and r^2^, for each fit for combinations of OBJECTID, HABITAT, and YEAR with at least `r min_samples` samples. In addition, the p-value of the fit is given. The figures also give the 99%-confidence and 99%-prediction intervals. See figure caption for more information.


```{r results='asis'}
# add points with confidence and prediction intervals
d$g <- d %>%
    select(-n) %>%
    pmap(function(OBJECTID, HABITAT, YEAR, x_logN, model, g) {
        g +
            geom_ribbon(
                data = data.frame(
                    logN = x_logN,
                    predict(model, newdata = data.frame(logN = x_logN), 
                           interval = "prediction", level = 0.99)
                ),
                mapping = aes(x = logN, ymin = lwr, ymax = upr),
                fill = "blue", alpha = 0.1) +
            geom_ribbon(
                data = data.frame(
                    logN = x_logN,
                    predict(model, newdata = data.frame(logN = x_logN), 
                           interval = "confidence", level = 0.99)
                ),
                mapping = aes(x = logN, ymin = lwr, ymax = upr),
                fill = "blue", alpha = 0.1) +
            geom_point(mapping = aes(x = logN, y = S-1)) +
            geom_path(
                data = data.frame(
                    logN = x_logN,
                    y = predict(model, newdata = data.frame(logN = x_logN))
                ),
                mapping = aes(x = logN, y = y), 
                colour = "blue"
            ) +
            scale_x_continuous(name = expression(ln(N)), limits = c(-0.026, NA)) +
            scale_y_continuous(name = "S-1") +
            ggtitle(sprintf("%s - %s - %s", OBJECTID, HABITAT, YEAR))
    }
)

# add r2
d$r2 <- d$model %>% 
    map_chr({. %>%
            summary %>%
            getElement("r.squared") %>%
            formatC(format = "f", digits = 2)
    })

# add fstats
d$f <- d$model %>%
    map({. %>% 
        summary %>%
        getElement("fstatistic")
    })

# add p-value
d$p <- d$f %>%
    map_chr(function(x) {
        if (is.null(x)) {
            return(NA_character_)
        } 
        pf( # see stats:::print.summary.lm
            q   = x[1L],
            df1 = x[2L],
            df2 = x[3L],
            lower.tail = FALSE
        ) %>%
        formatC(format = "f", digits = 5)
    })

d$g <- pmap(
    list(x = d$g, y = d$r2, z = d$p), 
    function(x, y, z) {
        x + 
            annotate("text", x = -Inf, y = Inf, 
                     label = paste0("r^2==", y), parse = TRUE, hjust = -0.1, vjust = 1.1) +
            annotate("text",  x = -Inf, y = Inf, 
                     label = paste0("p = ", z), hjust = -0.1, vjust = 3.1)
        }
    )

d$model %>%
    map_df(function(x) {
        x %>%
            coefficients %>%
            t %>%
            as_data_frame    
    }) %>%
    set_names(c("b0", "b1")) %>%
    prepend(d) %>%
    as_data_frame %>%
    select(OBJECTID, HABITAT, YEAR, `# samples` = n, b0, b1, r2, p) %>%
    arrange(OBJECTID, HABITAT, YEAR) %>%
    xtable %>% 
    print(type = "html")
```

<br>
<br>
<br>
<figure>
```{r,echo=FALSE, fig.retina=NULL, warning=FALSE, fig.width=4, fig.height=4, out.width=350, dpi=300, eval=isTRUE(nrow(d) > 0)}

d$g %>%
    walk(print)

```
<figcaption>Fits (blue line) of S-1 as function of N. The inner band is the 99%-confidence interval (represents our uncertainty about the fit), the outer band is the 99%-prediction interval (represents our uncertainty about individual points (future responses)).</figcaption>
</figure>
<br>
<br>

