
A plot of the `r title_text_plur` as function of `r settings$pressure$name` is given below, showing the sensitivity of each index. In addition, the `r settings$confidencelevel`-confidence interval of the mean `r title_text` is shown, indicating its precision.


<figure>
```{r,echo=FALSE, fig.retina=NULL, fig.width=5, fig.height=5, out.width=900, dpi=600, warning=FALSE, message=FALSE}

d <- d_ind %>% 
    select(OBJECTID, PRESSURE)

for (i in 1:nrow(W)) {
    w <- W[i, ]
    mmi_name <- paste0(names(w)[w > 0], collapse = "+")
    d[[mmi_name]] <- drop(V %*% w)
}

d <- d %>% 
    gather(key = "EQR", value = "VALUE", -OBJECTID, -PRESSURE)
d$EQR <- gsub(pattern = "_STAR", replacement = "*", x = d$EQR)

g <- ggplot(
    data = d, 
    mapping = aes(x = PRESSURE, y = VALUE, colour = EQR, linetype = EQR)) +
    stat_smooth(
        method = "lm", 
        formula = y ~ x, 
        level = settings$confidencelevel, 
        size = 1
    ) +
    scale_x_continuous(
        name = sprintf("%s (%s)", settings$pressure$name, settings$pressure$unit)) +
    scale_y_continuous(name = title_text) +
    scale_linetype_discrete(name = "") +
    scale_colour_discrete(name = "") +
    theme(axis.text = element_text(colour = "black"))
suppressWarnings(print(g))
```
<figcaption>Linear regression of `r title_text` as function of `r settings$pressure$name`. Data points are not shown. In addition, the `r settings$confidencelevel`-confidence interval of the `r title_text` is shown, indicating its precision. Note: plotting is restricted to the data range (no extrapolation).</figcaption>
</figure>
<br>
<br>



The plot below shows the same results, including the data points.


<figure>
```{r,echo=FALSE, fig.retina=NULL, fig.width=5, fig.height=5, out.width=700, dpi=600, warning=FALSE, message=FALSE}
for (eqr in unique(d$EQR)) {
    
    g <- ggplot(
            data = d %>% filter(EQR == eqr), 
            mapping = aes(x = PRESSURE, y = VALUE)
        ) +
        geom_point(alpha = 0.25) +
        stat_smooth(
            method = "lm", 
            formula = y ~ x, 
            level = settings$confidencelevel, 
            colour = "red"
        ) +
        scale_x_continuous(
            name = sprintf("%s (%s)", 
                           settings$pressure$name, 
                           settings$pressure$unit)) +
        scale_y_continuous(
            name = eqr, 
            limits = c(min(0, min(d$VALUE)), max(1, max(d$VALUE))),
            breaks = 0:5/5
        ) +
        theme(axis.text = element_text(colour = "black"))
    print(g)
}
```
<figcaption>`r Title_text` as function of `r settings$pressure$name`, including data points.</figcaption>
</figure>
<br>
<br>


Over-plotting may obscure relations in the data. Therefore, the plot below summarizes the dots in the plot above as (minimalistic) box plots:

- the blue outer (thinner) vertical line segment runs from the 5% to the 95% percentile;
- the blue inner (thicker) vertical line segment runs from the 25% to the 75% percentile;
- the dot is the median;
- the red line is the linear regression model fitted to the data by ordinary least squares.



<figure>
```{r,echo=FALSE, fig.retina=NULL, fig.width=5, fig.height=5, out.width=700, dpi=600, warning=FALSE, message=FALSE}
s <- d %>%
    group_by(OBJECTID, PRESSURE, EQR) %>%
    summarise(
        min = min(VALUE, na.rm = TRUE), 
        q05 = quantile(VALUE, probs = 0.05, na.rm = TRUE),
        q25 = quantile(VALUE, probs = 0.25, na.rm = TRUE),
        q50 = median(VALUE, na.rm = TRUE),
        q75 = quantile(VALUE, probs = 0.75, na.rm = TRUE),
        q95 = quantile(VALUE, probs = 0.95, na.rm = TRUE),
        max = max(VALUE, na.rm = TRUE),
        n = sum(!is.na(VALUE))
    )

for (eqr in unique(d$EQR)) {
    g <- ggplot(
            data = s %>% filter(EQR == eqr)
        ) +
        stat_smooth(
            data = d %>% filter(EQR == eqr),
            mapping = aes(x = PRESSURE, y = VALUE),
            method = "lm",
            formula = y ~ x,
            level = settings$confidencelevel,
            size = 0.5,
            colour = "red"
        ) +
        geom_linerange(
            mapping = aes(x = PRESSURE, ymin = q05, ymax = q95),
            size = 0.25,
            colour = "blue",
            alpha = 0.5
        ) +
        geom_linerange(
            mapping = aes(x = PRESSURE, ymin = q25, ymax = q75),
            size = 0.50,
            colour = "blue",
            alpha = 0.5
        ) +
        geom_point(
            mapping = aes(x = PRESSURE, y = q50)
        ) +
        scale_x_continuous(
            name = sprintf("%s (%s)", 
                           settings$pressure$name, 
                           settings$pressure$unit)) +
        scale_y_continuous(
            name = eqr, 
            limits = c(min(0, min(d$VALUE)), max(1, max(d$VALUE))),
            breaks = 0:5/5
        ) +
        theme(axis.text = element_text(colour = "black"))
    print(g)
}
```
<figcaption>`r Title_text` as function of `r settings$pressure$name`. See text for explanation.</figcaption>
</figure>
<br>
<br>
