### Box-whisker plots of Ecological Quality Ratio's

```{r}
to_log("INFO", "Entering subsection 'Box-whisker plots...'...")
```

The figures below show box-whisker plots of the `r title_text_plur` computed above for each OBJECTID.

The upper/lower hinge (_i.e._, the top and the bottom of the box respectively) corresponds to the third/first quartile. The upper/lower whisker extends from the third/first quartile hinge to the largest/smallest value no further than 1.5 times the inter-quartile range from that hinge. Extreme data values are plotted as points.

```{r}
fig_height <- ((nrow(W)-1L) %/% 3L + 1L) * 2.8
```
<figure>
```{r,echo=FALSE, fig.retina=NULL, fig.width=6, fig.height=fig_height, out.width=900, dpi=600, warning=FALSE, message=FALSE, eval=has_pressure}

d <- d_ind %>% 
    select(OBJECTID, PRESSURE)

for (i in 1:nrow(W)) {
    w <- W[i, ]
    mmi_name <- paste0(names(w)[w > 0], collapse = "+")
    mmi_name <- gsub(pattern = "_EQR", replacement = "", x = mmi_name)
    d[[mmi_name]] <- drop(V %*% w)
}
    
d <- d %>% 
    gather(key = "EQR", value = "VALUE", -OBJECTID, -PRESSURE)
d$EQR <- sub(pattern = "_STAR$", replacement = "*", x = d$EQR)

g <- ggplot(data = d, mapping = aes(x = OBJECTID, y = VALUE)) +
    geom_boxplot() +
    facet_wrap(~EQR) +
    scale_x_discrete(name = "") +
    scale_y_continuous(name = "") +
    theme(
        axis.text = element_text(colour = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1)
    )

suppressWarnings(print(g))
```
<figcaption>Box-whisker plots of the `r title_text_plur` for each OBJECTID</figcaption>
</figure>
<br>
<br>
