### Which MMI correlates best with pressure?

```{r}
to_log("INFO", "Entering subsection 'Which MMI correlates best with pressure?'...")
```


```{r}
number_of_combinations <- c(1L, 3L, 7L)[number_of_indicators]   
```

Multimetric indices are constructed by fitting all `r number_of_combinations` possible weighted linear combinations of the set of indices `r toString(paste0(indicators, "*"))` to the pressure. 

Our general model is

MMI^* = b<sub>0</sub> + b<sub>1</sub> &times; exp(b<sub>2</sub> &times; PRESSURE)

where MMI^* is the `r if (tolower(settings$legendtext) == "eqr") {"EQR of the"} else {"normalized"}` multimetric index, PRESSURE are the values in the PRESSURE column of the input file, and b<sub>0</sub>, b<sub>1</sub> and b<sub>2</sub> are coefficients.

MMI^* is a weighted linear combination of `r if (tolower(settings$legendtext) == "eqr") {"the ecological quality ratios of the"} else {"normalized values"}` of the selected indices:

`r paste(paste0("w<sub>", 1:number_of_indicators, "</sub>"), paste0(indicators, "*"), collapse = " + ", sep = " &times; ")`. 

where weights `r toString(paste0("w<sub>", 1:number_of_indicators, "</sub>"))` are non-negative and sum to one.

For each of the `r number_of_combinations` possible combinations of `r title_text_plur`, the weights are optimized mathematically in order to maximize the precision of the model mentioned above. Duplicated models (with approximately the same set of weights) have been removed.

```{r}
# initialize weight matrix and add single metric weights
W <- diag(number_of_indicators)
dimnames(W) <- list(NULL, indicators_EQR)
W <- rbind(W, 
    matrix(
        data = 0, 
        nrow = number_of_combinations - number_of_indicators, 
        ncol = number_of_indicators
    )
)



# bimetric indicators
i <- number_of_indicators
if (number_of_indicators >= 2L) {
    f_obj <- function(w, id) {
            (sprintf(
                "I(%e * %s + %e * %s) ~ b0 + b1*exp(b2*PRESSURE)",
                w,     indicators_EQR[id[1]],
                1 - w, indicators_EQR[id[2]]
            ) %>%
            as.formula %>%
            nls(
                data = d_ind,  
                start = list(b0 = 0.1, b1 = 0.9, b2 = -1),
                control = nls.control(maxiter = 1000))
            ) %>%
            AIC
    }
    opt <- optimize(f = f_obj, interval = c(0, 1), maximum = FALSE, id = c(1, 2))
    i <- i + 1L
    W[i, 1] <-     opt$minimum
    W[i, 2] <- 1 - opt$minimum
}


# trimetric indicators
if (number_of_indicators == 3L) {
    opt <- optimize(f = f_obj, interval = c(0, 1), maximum = FALSE, id = c(1, 3))
    i <- i + 1L
    W[i, 1] <-     opt$minimum
    W[i, 3] <- 1 - opt$minimum
    opt <- optimize(f = f_obj, interval = c(0, 1), maximum = FALSE, id = c(2, 3))
    i <- i + 1L
    W[i, 2] <-     opt$minimum
    W[i, 3] <- 1 - opt$minimum
}

if (number_of_indicators == 3L) {
    f_obj <- function(w) {
        if (any(w < 0)) {
            return(Inf)
        }
        w <- w / sum(w)
        (sprintf(
            "I(%e * %s + %e * %s + %e * %s) ~ b0 + b1*exp(b2*PRESSURE)",
            w[1], indicators_EQR[1],
            w[2], indicators_EQR[2],
            w[3], indicators_EQR[3]
        ) %>% 
        as.formula %>%
        nls(
            data = d_ind,  
            start = list(b0 = 0.1, b1 = 0.9, b2 = -1),
            control = nls.control(maxiter = 1000))
        ) %>%
        AIC
    }
    opt <- optim(
        par = c(1, 1, 1), 
        fn = f_obj, 
        method = "Nelder-Mead", 
        control = list(maxit = 1000, fnscale = 1) # minimization
    )
    has_converged <- opt$convergence == 0L
    if (!has_converged) {
        stop("Can't find an optimum set of weights", call. = FALSE)
    }
    i <- i + 1L
    W[i, ] <- opt$par / sum(opt$par)
}
```
